Discovering cell type motifs in connectomics data

### 2013-12-03 docs/paper.tex
so happy with this progress. Starting to feel like a real body of work.

\documentclass{nature}
\usepackage[english,american]{babel}
\usepackage{subfiles}
\usepackage{grffile}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{fullpage}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{amsmath}
\usepackage{subcaption}
\usepackage{multicol}
\usepackage[citestyle=authoryear,url=false,hyperref=auto,isbn=false]{biblatex}
\addbibresource{connectomics motif.bib}
\addbibresource{connectomics motif-relational models.bib}

@@

\begin{figure}
  \centering 
  \centerline{\includegraphics[width=8.5in]{overview.ai}}
  \caption{This is the pedagogy figure where we explain what is is
    that the model does. }
\end{figure}

@@

\begin{figure}
  \centering 
  \centerline{\includegraphics[width=8.5in]{synthetic.ai}}
  \caption{This is the synthetic data figure where we show that our model works with synthetic data across a variety of topologies and geometries}
\end{figure}

@@

\begin{figure}
  \centering 
  \centerline{\includegraphics[width=8.5in]{mouseretina.ai}}
  \caption{Mouse retina data}
\end{figure}

@@

\begin{figure}
  \centering 
  \centerline{\includegraphics[width=8.5in]{celegans.ai}}
  \caption{C elegans data}
\end{figure}

@@

\begin{figure}
  \centering 
  \centerline{\includegraphics[width=8.5in]{mos6502.ai}}
  \caption{Microprocessor data}
\end{figure}

@@

We then jointly compute the maximum a posteriroi estimate of the class
assignment vector ${c_i}$, the parameter matrix $\eta_{mn}$, and the
global model hyperparameters $\theta$ :

@@

We describe three link functions, ``Logistic-distance'',

Prior on cluster assignment. While dirichlet process is inconsistent,
we deemed this to ultimately be inconsequential for this study.

Priors on gridded hyperparameter space.

@@

Our first transition kernel (``structural'') performs gibbs sampling 
of the assignment vector $p(\vec{c} | \eta, \theta, \alpha)$. 
The lack of conjugacy in our likelihood model makes an explicit 
evaluation of the conditional assignment probabilities impossible, 
motivating us to use an auxiliary variable method \autocite{Neal}
in which a collection of ephemeral clusters are explicitly represented
for the duration of the Gibbs scan.

We then employ a transition kernel to update the per-component
parameter values $\eta_{mn}$. Conditioned on the assignment vector
$\vec{c}$ and the model hyperparameters $\theta, \alpha$ the 
individual $\eta_{mn}$ are indepednent. We slice sample \autocite{neal}
each component's parameters, choosing the slice width as a function
of the global hyperparameter range.

@@

\begin{addendum}
 \item Put acknowledgements here.
 \item[Competing Interests] The authors declare that they have no
competing financial interests.
 \item[Correspondence] Correspondence and requests for materials
should be addressed to A.B.C.~(email: myaddress@nowhere.edu).
\end{addendum}

>>>>

%\documentclass{nature}
\documentclass{article}
\usepackage[english,american]{babel}
\usepackage[backend=biber,style=nature]{biblatex}
\usepackage{subfiles}
\usepackage{grffile}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{fullpage}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{amsmath}
\usepackage{subcaption}
\usepackage{multicol}

\addbibresource{connectomics motif.bib}
\addbibresource{connectomics motif-relational models.bib}

@@

\begin{figure}
  \centering 
    \centerline{\includegraphics[width=183mm]{overview.ai}}
  \caption{An overview of our method. a. Input data is
an adjacency matrix indicating whether cell $i$ (row)
synapses on cell $j$ (column). The location of each cell in space is known. 
b. Our algorithm discovers hidden cell types in this connectivity data
by assuming all cells of a type share a distance-dependent connectivity 
profile, c.), with cells of other types. This results in a clustering 
of the cells by those hidden types. b.) shows the cell adjacency
matrix with cells of the same type grouped together. d.) shows
the learned probability of connection between our different types
at various distances, illustrating short-range and long-range
connectivtiy patterns.}

\end{figure}

@@

\begin{figure}
  \centering 
  \centerline{\includegraphics{synthetic.ai}}
  \caption{Correct recovery of true numbers of hidden classes
in synthetic data. a.) Other models overcluster the number
of classes as they fail to take distance into account, whereas
our model finds close to the true number of classes. b.) ARI, a measure
of ``correct clustering'', for
different true class counts, between our model and a traditional
nonparametric stochastic block model. c.) Multiple Markov chains iterateively
converge towards a highly-probable solution. }

\end{figure}

@@

\begin{figure}
  \centering 
  \centerline{\includegraphics{mouseretina.ai}}
  \caption{Discovering cell classes in the mouse retina connectome. 
a.) Input adajacency data for 950 cells for which soma positions were known. b. clsutered adjacency matrix. c.) connectivity diagram showing our clusters, as
well as the cell depth and known cell type of each cell. Our model agrees
with the cell types traditionally identified by anatomists, and identifies 
laminar-specific connectivity patterns. d.) Connectivity between our
clusters as a function of distance -- the cluster consisting primarially of
retinal ganglion cells (lower left-center) exhibits the expected near and
far connectivity.}
\end{figure}

@@

\begin{figure}
  \centering 
  \centerline{\includegraphics{celegans.ai}}
  \caption{Discovering cell classes in the \textit{c. elegans} connectome. a.) Initial cell connectivity matrix, red are (directional) chemical synapses, blue are
electrical synapses. Point size is proportaionl to synapse count between those cells. b.) Clustered adjaceny matrix showing identified clusters. c.) Inter-cluster
connectivity, showing the position of each cell along the body axis as well
as the ``true'' cell type -- both coarse (motor, sensory, and interneuron) as well
as fine (small text labels). }
\end{figure}

@@

\begin{figure}
  \centering 
  \centerline{\includegraphics{mos6502.ai}}
  \caption{Microprocessor data}
\end{figure}

@@

We then jointly compute the maximum a posteriroi (MAP) estimate of the
class assignment vector ${c_i}$, the parameter matrix $\eta_{mn}$, and
the global model hyperparameters $\theta$ :

@@

We describe three link functions, ``Logistic-distance Bernoulli'',
``logistic-distance fixed-lambda Bernoulli'', and ``logistid-Distance
Poisson''.

The ``logistic-distance Bernoulli'' model associates with each latent class 
two parameters, $\mu_{mn}$ and $\lambda_{mn}$, and the probability that two
cells $e_i$ and $e_j$ are connected is given by
\begin{equation}
p^* = \frac{1.0}{1 + \exp \frac{d(e_i, e_j) - \mu_{mn}}{\lambda_{mn}}}\\
p = p^* \cdot (p_{max} - p_{min}) + p_{min}
\end{equation}
where $p_{max}$ and $p_{min}$ are per-graph hyperparameters. Per-component parameters $\mu_{mn} \sim \exp(\mu | \mu^{hp})$ and $\lambda_{mn} \sim \exp(\lambda | \lambda^{hp})$. 
Inference is performed on these hyperparameters as well, see TODO supplemental 
for details.

We use a Dirichlet-process prior on cluster assignments, which allows
the number of clusters to be determined automatically. In brief, the
probability of a cell belonging to a cluster is proportional to the
number of datapoints already in that cluster, $m_k$, such that $p(c_i
= k) \propto \frac{m_k}{N + \alpha}$ and the probability of the cell
belonging to a new cluster $k'$ is $p(c_i = k') \propto
\frac{\alpha}{N + \alpha}$.

@@

Our first transition kernel (``structural'') performs gibbs sampling 
of the assignment vector $p(\vec{c} | \eta, \theta, \alpha)$. 
The lack of conjugacy in our likelihood model makes an explicit 
evaluation of the conditional assignment probabilities impossible, 
motivating us to use an auxiliary variable method \autocite{Neal2000}
in which a collection of ephemeral clusters are explicitly represented
for the duration of the Gibbs scan.

We then employ a transition kernel to update the per-component
parameter values $\eta_{mn}$. Conditioned on the assignment vector
$\vec{c}$ and the model hyperparameters $\theta, \alpha$ the 
individual $\eta_{mn}$ are indepednent. We slice sample \autocite{NealSlice}
each component's parameters, choosing the slice width as a function
of the global hyperparameter range.

@@

% \begin{addendum}
%  \item Put acknowledgements here.
%  \item[Competing Interests] The authors declare that they have no
% competing financial interests.
%  \item[Correspondence] Correspondence and requests for materials
% should be addressed to A.B.C.~(email: myaddress@nowhere.edu).
% \end{addendum}

\printbibliography
